---
title: "Untitled"
format: pdf
editor: visual
---

```{r}
install.packages("tidyverse")
```


# Loading required libraries

```{r}
library(tidyverse)
library(readxl)
```

# Downloading the data

```{r}
download.file("https://raw.githubusercontent.com/toryn13//Rexamples/main/IL6.xlsx", "IL6.xlsx")
download.file("https://raw.githubusercontent.com/toryn13//Rexamples/main/IL6plate.xls", "IL6plate.xls")
```

# Reading and manipulating data

```{r}
elisa_data <- read_excel("IL6plate.xls", skip = 1, n_max = 9)
il6data <- read_excel("IL6plate.xls", range = "D2:E10")
colnames(il6data) <- c("Control", "PHA")
standards <- read_excel("IL6plate.xls", range = "A14:D22")
colnames(standards) <- c("cell", "Conc", "rep1", "rep2")
standardsL <- pivot_longer(standards, starts_with("rep"), names_to = "replicate", values_to = "Abs")
```

# Plotting the standard curve

```{r}
curve <- standardsL %>% ggplot(aes(Conc, Abs)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_cowplot(font_size = 20, line_size = 1) + 
  xlab("IL6 pg/mL")
curve
```

# Calculating mean Abs values for standards

```{r}
mean_standards <- standardsL %>% group_by(Conc) %>% summarise(meanAbs = mean(Abs))
calc <- lm(meanAbs ~ Conc, data = mean_standards)
```

# Mutating IL6 data for concentration calculation

```{r}
il6data <- il6data %>% mutate(Concon = (Control - 0.1693065) / 0.0050953, PHAconc = (PHA - 0.1693065) / 0.0050953)
```

# Reading the IL6 treatment data and adjusting columns

```{r}
IL6 <- read_excel("IL6.xlsx")
IL6 <- IL6 %>% mutate(Treat2 = case_when(Treat == "Control" ~ "1", Treat == "PHA" ~ "2"))
IL6$Conc <- (IL6$Conc) / 1000
IL6$Treat2 <- as.numeric(IL6$Treat2)
IL6$Treat2 <- factor(IL6$Treat2)
IL6$Day <- factor(IL6$Day)
```

# Summarizing the IL6 data

```{r}
sumIl6 <- IL6 %>% group_by(Treat, Day) %>% summarise(meanIL6 = mean(Conc), sdIL6 = sd(Conc))
sumIl6$Day <- factor(sumIl6$Day)
```

# Plotting the results

```{r}
p2 <- sumIl6 %>% ggplot(aes(Day, meanIL6, fill = Treat)) + 
  geom_bar(stat = "identity", position = position_dodge(), colour = "black") + 
  geom_errorbar(aes(ymin = meanIL6 - sdIL6, ymax = meanIL6 + sdIL6), position = position_dodge(.9), width = 0.6) + 
  scale_y_continuous(limits = c(0, 50), expand = c(0, 0)) + 
  scale_fill_manual(values = c('#66b3ff', '#cce6ff')) + 
  theme_cowplot(font_size = 20) + 
  ylab("mean IL6 ng/mL")
p2
```

# T-tests and filtering

```{r}
IL6$group <- factor(paste0(IL6$Treat, "_", IL6$Day))
t.test(Conc ~ Treat, data = IL6)
pha <- IL6 %>% filter(group %in% c("PHA_1", "PHA_3"))
t.test(Conc ~ group, data = pha)
```

# ANOVA and Tukey's post hoc test

library(rstatix)

```{r}
IL6 %>% 
  t_test(data = ., Conc ~ group) %>% 
  adjust_pvalue(method = "bonferroni") %>% 
  add_significance("p.adj")

# One-way ANOVA with post hoc test
library(broom)
IL6.aov <- aov(Conc ~ group, data = IL6)
tidy(IL6.aov)
pht <- data.frame(tidy(TukeyHSD(IL6.aov, which = "group")))
write.csv(pht, "pht.csv")
```

```{r}
# Two-way ANOVA
IL6.aov2 <- aov(Conc ~ Day + Treat, data = IL6)
tidy(IL6.aov2)
IL6.aov2 <- aov(Conc ~ Day * Treat, data = IL6)
tidy(IL6.aov2)
```

# Linear models

```{r}
aLM <- lm(Conc ~ group, data = IL6)
tidy(aLM)
a2LM <- lm(Conc ~ Day + Treat, data = IL6)
tidy(a2LM)
```

# Plotting with GGPLOT2

```{r}
sumIl6 %>% ggplot(aes(Day, meanIL6, fill = Treat)) + geom_bar(stat = "identity")
sumIl6 %>% ggplot(aes(Day, meanIL6, fill = Treat)) + geom_bar(stat = "identity", position = position_dodge())
```

# Error bars

```{r}
sumIl6 %>% ggplot(aes(Day, meanIL6, fill = Treat)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = meanIL6 - sdIL6, ymax = meanIL6 + sdIL6), width = .1) + 
  scale_y_continuous(limits = c(0, 50), expand = c(0, 0))

sumIl6 %>% ggplot(aes(Day, meanIL6, fill = Treat)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = meanIL6 - sdIL6, ymax = meanIL6 + sdIL6), position = position_dodge(.9), width = 0.8) + 
  scale_y_continuous(limits = c(0, 50), expand = c(0, 0))

p2 <- sumIl6 %>% ggplot(aes(Day, meanIL6, fill = Treat)) + 
  geom_bar(stat = "identity", position = position_dodge(), colour = "black") + 
  geom_errorbar(aes(ymin = meanIL6 - sdIL6, ymax = meanIL6 + sdIL6), position = position_dodge(.9), width = 0.6) + 
  scale_y_continuous(limits = c(0, 50), expand = c(0, 0)) + 
  scale_fill_manual(values = c('#66b3ff', '#cce6ff')) + 
  theme_bw() + 
  ylab("mean IL6 ng/mL")
p2
```
