---
title: "Untitled"
output: html_document
date: '2022-07-06'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Free data science book https://r4ds.had.co.nz/

# Fetch the files
```{r}
download.file("https://raw.githubusercontent.com/toryn13//Rexamples/main/cat.jpg", "cat.jpg")
download.file("https://raw.githubusercontent.com/toryn13//Rexamples/main/IL6.xlsx", "IL6.xlsx")
download.file("https://raw.githubusercontent.com/toryn13//Rexamples/main/IL6plate.xls", "IL6plate.xls")
download.file("https://raw.githubusercontent.com/toryn13//Rexamples/main/p53.jpeg", "p53.jepg")
```

# Here are the packages we need 
```{r}
install.packages("cowplot")
install.packages("rstatix")
install.packages("magick")
install.packages("tidyverse")
```


# Load the packages
```{r}
library(tidyverse)
library(cowplot)
library(magick)
library(readxl)
library(rstatix)
```


<- assignment


%>% pipe

log2(25)

25%>%log2()

#Importing data

We're using read_excel, if you have a tsv, csv or text. Use the readr package

Excel sheets or regions of cells can be imported.

R has many data structures. These include

- vector
- list
- matrix
- data frame
- factors


```{r}
elisa_data<-read_excel("IL6plate.xls", skip=1,n_max = 9)
```

```{r}
standards<-read_excel("IL6plate.xls", range = "A14:D22")
colnames(standards)<-c("cell", "Conc", "rep1", "rep2")
```

5 common data types.

- character
- numeric (real or decimal)
- integer
- logical
- complex


#Wide vs long format
tidy data
```{r}
standardsL<-pivot_longer(standards, starts_with("rep"), names_to = "replicate", values_to = "Abs")
```




#GGLOT2, building plots in layers
```{r}

standardsL%>%ggplot(aes(Conc,Abs))


standardsL%>%ggplot(aes(Conc,Abs))+geom_point()


standardsL%>%ggplot(aes(Conc,Abs))+geom_point()+geom_line()


standardsL%>%ggplot(aes(Conc,Abs, group=replicate))+geom_point(colour="blue")+geom_line(colour="red")


standardsL%>%ggplot(aes(Conc,Abs, group=replicate))+geom_point(colour="blue")+geom_line(colour="red")+facet_grid(.~replicate)


standardsL%>%ggplot(aes(Conc,Abs, color=replicate))+geom_point()+geom_line()+facet_grid(.~replicate)


curve<-standardsL%>%ggplot(aes(Conc,Abs))+geom_point()+geom_smooth(method="lm")+theme_cowplot(font_size=20, line_size = 1)+xlab(c("IL6 pg/mL"))
curve
```

#Modifying values and creating new variables
```{r}
IL6<-read_excel("IL6.xlsx")
IL6<-IL6%>%mutate(Treat2 = case_when(Treat=="Control"~"C", Treat=="PHA"~"P"))
IL6$Conc<-(IL6$Conc)/1000
```
#Factors
```{r}
IL6$Treat2<-as.numeric(IL6$Treat2)
IL6$Treat2<-factor(IL6$Treat2)
IL6$Day<-factor(IL6$Day)
```
#boxplots
```{r}
IL6%>%ggplot(aes(Treat2, Conc,fill=Treat))+geom_boxplot()+scale_fill_manual(values=c('blue', 'lightblue'))+theme_bw()


IL6%>%ggplot(aes(Treat2, Conc,fill=Treat))+geom_boxplot()+scale_fill_manual(values=c('blue', 'lightblue'))+theme_cowplot(12)


IL6%>%ggplot(aes(Treat, Conc,fill=Treat))+geom_boxplot()+geom_jitter()+scale_fill_manual(values=c('blue', 'lightblue'))+theme_cowplot(font_size=20)

```



```{r}
boxplot<-IL6%>%ggplot(aes(Treat, Conc,fill=Treat))+geom_boxplot()+geom_jitter(size=3)+
scale_fill_manual(values=c('blue', 'lightblue'))+
theme_cowplot(font_size=20, line_size=1)+theme(legend.position='none')+
xlab(NULL)+ylab("IL6 ng/mL")+
scale_y_continuous(limits = c(0, 50), expand = c(0, 0))
boxplot
```

#Summary data
```{r}
sumIl6<-IL6%>%group_by(Treat,Day)%>%summarise(meanIL6=mean(Conc), sdIL6=sd(Conc))
sumIl6
```

```{r}
sumIl6$Day<-factor(sumIl6$Day)
```
#Bar plot
```{r}
#sumIl6%>%ggplot(aes(Day, meanIL6, fill=Treat))+geom_bar()
sumIl6%>%ggplot(aes(Day, meanIL6, fill=Treat))+geom_bar(stat="identity")
sumIl6%>%ggplot(aes(Day, meanIL6, fill=Treat))+geom_bar(stat="identity",position = position_dodge())
```

#Bar plot with error bars
```{r}
sumIl6%>%ggplot(aes(Day, meanIL6, fill=Treat))+geom_bar(stat="identity", position = position_dodge())+
geom_errorbar(aes(ymin = meanIL6-sdIL6, ymax = meanIL6+sdIL6), width=.1)+
scale_y_continuous(limits = c(0, 50), expand = c(0, 0))
````


```{r}
sumIl6%>%ggplot(aes(Day, meanIL6, fill=Treat))+geom_bar(stat="identity", position = position_dodge())+
geom_errorbar(aes(ymin = meanIL6-sdIL6, ymax = meanIL6+sdIL6),position=position_dodge(.9), width=0.8)+
scale_y_continuous(limits = c(0, 50), expand = c(0, 0))
```

```{r}
p2<-sumIl6%>%ggplot(aes(Day, meanIL6, fill=Treat))+geom_bar(stat="identity", position = position_dodge(),colour="black")+
geom_errorbar(aes(ymin = meanIL6-sdIL6, ymax = meanIL6+sdIL6),position=position_dodge(.9), width=0.6)+
scale_y_continuous(limits = c(0, 50), expand = c(0, 0))+
scale_fill_manual(values=c('#66b3ff','#cce6ff'))+theme_cowplot(font_size = 20)+
ylab("mean IL6 ng/mL")
p2
```

#add the orginal data points
```{r}
p3<-p2+geom_dotplot(data=IL6, aes(Day, Conc, fill=Treat),binaxis = "y",stackdir = "center", binpositions="all", position="dodge" )
p3
```


#Many other plot types
```{r}
heat_map <- IL6%>%ggplot(aes(x = Day,y = Treat, fill = scale(Conc))) + geom_tile()+
scale_fill_gradient2(low = "blue", mid="white", high = "red")+coord_flip()+theme_minimal_grid()
heat_map
````



# Statistics
```{r}
IL6$group<-factor(paste0(IL6$Treat,"_",IL6$Day))
str(IL6)
```

```{r}
IL6%>%t_test(Conc~Treat)
```

#Using filter
```{r}
IL6%>%filter(group=="Control_1")
IL6%>%filter(group!="Control_1")
IL6%>%filter(group %in% c("PHA_1", "PHA_3"))
pha<-IL6%>%filter(group %in% c("PHA_1", "PHA_3"))
```

pha

```{r}
IL6 %>%
t_test(data =., Conc ~ group) %>%
adjust_pvalue(method = "bonferroni") %>%
add_significance("p.adj")
```

#Annotating plots
```{r}
bar<-p3+annotate("segment", x = 1.2, xend = 2.3, y = 45,yend=45,
  colour = "black")+annotate("text",x = 1.8, y = 46, label="***",size=10)
bar
```

# Images 
```{r}
cat<-image_read("cat.jpg")
#format width height 
#JPEG     389    428
#crop image Crop image to a by b at position c,d
crop_cat<-image_crop(cat, "200x200+100+100")

#scale to width
small_cat<-image_scale(cat, "200")
small_crop_cat<-image_scale(crop_cat, "100")


image_write(small_cat, path = "small_cat.png", format = "png")
image_write(crop_cat, path = "crop_cat.png", format = "png")
image_write(small_crop_cat, path = "small_crop_cat.png", format = "png")
```

# Example with western blot
```{r}
a<-image_read("p53.jepg")
print(a)
a<-image_convert(a, "png")
a<-image_crop(a, "150x250+220+350")
```

```{r}
a_text<-image_annotate(a, "1      2", size = 35, gravity = "North", color = "black")
a_text<-image_annotate(a_text, "Crop from p53 blot", size = 12, gravity = "south", color = "black")

image_write(a_text, path = "crop.png", format = "png")
```

```{r}
blot<-ggdraw() +
draw_image("crop.png")
```

# Adding images to GGPLOT2
```{r}
C1=seq(1:10)
C2=seq(1:10)
frame<-data.frame(C1,C2)
```

frame

```{r}
graph<-frame%>%ggplot(aes(C1, C2))+geom_point(color="white")+theme_nothing()+annotate("segment", x = 0, xend = 0.9, y =  5.5,yend=5.5,
  colour = "black", arrow=arrow())+annotate("segment", x = 0, xend = 0.9, y =  7.5,yend=7.5,
  colour = "black", arrow=arrow())+annotate("text",x = 9, y = 5.5, label="p53",size=10)
```

graph

```{r}

img<-a_text%>%image_draw()
````

```{r}
blot2<-ggdraw()+draw_plot(graph)+draw_image(img)
blot2
```

```{r}
plot_grid(curve,bar, labels = "AUTO")
ggsave("example.pdf")


plot_grid(blot2,boxplot, labels = "AUTO", scale = c(0.7, 1))
ggsave("example2.pdf")
```





